---
title: "Project title"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: true
---

**Introduction and Data:**

The data comes from sports-reference.com. The data was collected five years and was scraped from sports-reference.com by Randi Griffin, an Evolutionary Anthropology graduate student of Duke University. Randi also competed in the 2018 Winter Olympics on the Unified Korea women's national ice hockey team. This was initially made into a visualization that was then further adapted to the current data set.

The observations represent every athlete who has won a medal at the Winter Olympics. The values in each observation include the year of the Olympics, the type of event the medal was won in, the type of medal won, the country the Olympian was playing for, the gender of the Olympian, and the name of the Olympian. Therefore, the data does include both individual medal winners as well as winning teams, and also includes medalists from countries that no longer exist.

A possible ethical issue with this dataset is that the Olympics page on sports-reference.com has since been removed from the website. Due to low traffic on the Olympics page, less work was done to check over the information and verify that it is all correct, and then sports-reference decided it was just better to take down the data. Because of this, a few of the names of the athletes are slightly misspelled or have extraneous characters, and the page was not updated with results from the 2018 or 2022 Winter Olympics. Despite this, the data was collected ethically, sports-reference is cited as the source of the data, and all of the Terms of Use listed on the sports-reference website appear to have been followed when the data was taken from sports-reference.com. 

\

**Methodology:**

Our research question is "What are the factors that influence Winter Olympic success across countries, events and year it took place in?"

Our research topic focuses on analyzing the performance of athletes who have won medals at the Winter Olympics. We plan to answer this research question through interference and not prediction, since we are analyzing past data that is concrete. Therefore, we do not need to predict outcomes of data since we are just inferring patterns and trends based on existing data. We plan to gather results from our data by creating visuals comparing/contrasting factors such as gender, age, Olympic event and more to see which factors affected Olympic winter success. 

This research will identify patterns and trends in medal-winning performances across different countries, events, and time periods. We hypothesize that more cold-weathered countries, populous countries, and younger participants will achieve more wins each year in the Winter Olympics. 

The variables represented are all categorical variables: countries, athletes, events, type of medals, and years.

**Adding Additional Variables**

```{r}
#|: category

WinterOlympic <- WinterOlympic |>
  mutate(category = case_when(str_detect(Sport, "Alpine Skiing") ~ "Skiing/Snowboarding", 
                              str_detect(Sport, "Biathlon") ~ "Nordic", 
                              str_detect(Sport, "Bobsled") ~ "Ice", 
                              str_detect(Sport, "Cross-Country Skiing") ~ "Nordic",
                              str_detect(Sport, "Curling") ~ "Ice", 
                              str_detect(Sport, "Figure Skating") ~ "Ice", 
                              str_detect(Sport, "Freestyle Skiing") ~ "Skiing/Snowboarding",
                              str_detect(Sport, "Ice Hockey") ~ "Ice", 
                              str_detect(Sport, "Luge") ~ "Ice", 
                              str_detect(Sport, "Nordic Combined") ~ "Nordic", 
                              str_detect(Sport, "Short-Track Speedskating") ~ "Ice", 
                              str_detect(Sport, "Skeleton") ~ "Ice", 
                              str_detect(Sport, "Ski Jumping") ~ "Nordic", 
                              str_detect(Sport, "Snowboarding") ~ "Skiing/Snowboarding", 
                              str_detect(Sport, "Speedskating") ~ "Ice"))

WinterOlympic |>
  count(category)

```

**Results**

**Coding Ideas**

What is the relationship between age and medals won in specific Winter Olympics events? Demonstrate using a visualization.

```{r}
#| label: Create-Visualization
```

Process: 1. Group the Winter Olympic events into similar categories (ice events, skiing/snowboarding events, Nordic events) 2. Filter out team events because there is no age variable 3. Created faceted bar plot with grouped events.

What is the relationship between age and quantity of medals won in the Winter Olympics? Demonstrate using a visualization.

```{r}
#| label: Create-Visualization-2

```

Process: 1. 2. 3.

Given that a participant was from a Scandinavian country, what is the probability that they won a gold, silver, or bronze medal in a Nordic event, as these originated in Scandinavia?

```{r}
#| label: Calculate-Probability

Scandinavia <- WinterOlympic |>
  filter(Country == "Norway" | 
         Country == "Sweden" | 
         Country == "Denmark" | 
         Country == "Finland")

Scandinavia |>
  count(category) |>
  mutate(prob = n / nrow(Scandinavia)) |>
  filter(category == "Nordic")

WinterOlympic |>
  count(category) |>
  mutate(prob = n / nrow(WinterOlympic)) |>
  filter(category == "Nordic")

```

The probability that a medal won by a Scandinavian athlete was in a Nordic event is 0.644.

The probability that any medal won in the Winter Olympics was from a Nordic event is 0.326.

Process: 1. 2. 3.

\

\
\
