---
title: "Project title"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: true
---

**Introduction and Data:**

The data comes from sports-reference.com. The data was collected five years and was scraped from sports-reference.com by Randi Griffin, an Evolutionary Anthropology graduate student of Duke University. Randi also competed in the 2018 Winter Olympics on the Unified Korea women's national ice hockey team. This was initially made into a visualization that was then further adapted to the current data set.

The observations represent every athlete who has won a medal at the Winter Olympics. The values in each observation include the year of the Olympics, the type of event the medal was won in, the type of medal won, the country the Olympian was playing for, the gender of the Olympian, and the name of the Olympian.

**Methodology:**

Our research question is "What are the factors that influence Winter Olympic success across countries, events and year it took place in?"

Our research topic focuses on analyzing the performance of athletes who have won medals at the Winter Olympics. We plan to answer this research question through interference and not prediction, since we are analyzing past data that is concrete. Therefore, we do not need to predict outcomes of data since we are just inferring patterns and trends based on existing data. We plan to gather results from our data by creating visuals comparing/contrasting factors such as gender, age, Olympic event and more to see which factors affected Olympic winter success. 

This research will identify patterns and trends in medal-winning performances across different countries, events, and time periods. We hypothesize that more cold-weathered countries, populous countries, and younger participants will achieve more wins each year in the Winter Olympics. 

The variables represented are all categorical variables: countries, athletes, events, type of medals, and years.

**Adding Additional Variables**

**Results**

**Coding Ideas**

What is the relationship between age and medals won in specific Winter Olympics events? Demonstrate using a visualization.

```{r}
#| label: Create-Visualization
```

Process: 1. Group the Winter Olympic events into similar categories (skating, skiing/snowboarding, team events, track events) 2. Filter out team events because there is no age variable 3. Created faceted bar plot with grouped events.

<<<<<<< HEAD
What is the relationship between age and quantity of medals won in the Winter Olympics? Demonstrate using a visualization.
=======
Do men and women from from the same countries have similar success in Olympic Events?
>>>>>>> 1658718519e495f6c5e942c16ed5cfe323d0e523

```{r}
#| label: Create-Visualization-2

```

<<<<<<< HEAD
Process: 1. 2. 3.
=======
Process: 1. Filter out mixed events 2. Some sort of groupby for event 3. Bar graph, each bar is different country and stack separating color by gender (blue and pink?)
>>>>>>> 1658718519e495f6c5e942c16ed5cfe323d0e523

Given that a participant was from a Scandinavian country, what is the probability that they won a gold, silver, or bronze medal in a skiing event?

```{r}
#| label: Calculate-Probability
```

Process: 1. 2. 3.

\

\
\
